---
title: "Linear Regression with R"
format:
  html:
    code-tools:
      source: repo
---

```{r setup}
#| include: false
library(tidyverse)
knitr::opts_chunk$set(comment = NA)
html_table <- function(x, page_length = nrow(x)) {
  DT::datatable(
    data = x, 
    rownames = FALSE, 
    options = list(
      dom = 'tip', 
      scrollX = TRUE, 
      pageLength = page_length
    )
  )
}
dt_glance <- function(x) {
  x |> 
    broom::glance() |> 
    html_table() |> 
    DT::formatSignif(columns = c(1:5, 7:10), digits = 3)
}
dt_tidy <- function(x) {
  x |> 
    broom::tidy() |> 
    html_table() |> 
    DT::formatSignif(columns = 2:5, digits = 3)
}
```

## Expectations

In the following exercises, you will be asked to study the relationship of a
continuous response variable and one or more predictors. In doing so, remember
to:

- perform model diagnosis
- including visualization tools
- including multicollinearity assessment
- perform informed model selection
- comment each result of an analysis you run with R

## Exercise 1

Analysis of the `production` data set which is composed of the following
variables:

Variable name | Description
------------- | -------------
x             | Number of produced pieces
y             | Production cost

Study the relationship between `x` and `y`.

## Exercise 2

Analysis of the `brain` data set which is composed of the following variables:

Variable name | Description
------------- | -------------
body_weight   | Body weight in kg
brain_weight  | Brain weight in kg

Study the relationship between body and brain weights, to establish how the variable `brain_weight` changes with the variable `body_weight`.

## Exercise 3

Analysis of the `anscombe` data set which is composed of the following variables:

Variable name | Description
------------- | -------------
x1            | Predictor to be used for explaining `y1`
x2            | Predictor to be used for explaining `y2`
x3            | Predictor to be used for explaining `y3`
x4            | Predictor to be used for explaining `y4`
y1            | Response to be explained by `x1`
y2            | Response to be explained by `x2`
y3            | Response to be explained by `x3`
y4            | Response to be explained by `x4`

Study the relationship between each $y_i$ and the corresponding $x_i$. 

```{r anscombe}

```

## Exercise 4

Analysis of the `cement` data set, which contains the following variables:

Variable name     | Description
----------------- | ----------------
aluminium         | Percentage of $\mathrm{Ca}_3 \mathrm{Al}_2 \mathrm{O}_6$
silicate          | Percentage of $\mathrm{C}_2 \mathrm{S}$
aluminium_ferrite | Percentage of $4 \mathrm{CaO} \mathrm{Al}_2 \mathrm{O}_3 \mathrm{Fe}_2 \mathrm{O}_3$
silicate_bic      | Percentage of $\mathrm{C}_3 \mathrm{S}$
hardness          | Hardness of the cement obtained by mixing the above four components

Study, using a multiple linear regression model, how the variable `hardness` depends on the four predictors.

## Exercise 5

Analysis of the `job` data set, which contains the following variables:

Variable name | Description
------------- | ----------------
average_score | Average score obtained by the employee in the test
years_service | Number of years of service
sex           | Male or female

We want to see if it is possible to use the sex of the person in addition to the years of service to predict, with a linear model, the average score obtained in the test. Estimate a linear regression of `average_score` vs. `years_service`, considering the categorical variable `sex`.

## Exercise 6

Analysis of the `cars` data set, which contains the following variables:

Variable name | Description
------------- | ----------------
speed         | Speed of the car before starting braking
dist          | Distance travelled by the car during the braking period until it completely stops

Verify if the distance travelled during the braking depends on the starting velocity of the car:

1. Choose the best model to explain the distance as function of the speed,
2. Predict the braking distance for a starting velocity of 25 km/h, using a point estimate and a prediction interval.

```{r cars}

```

## Exercise 7

Analysis of the `mussels` data set, which contains the following variables:

Variable name | Description
------------- | ----------------
length        | Length of a mussel (mm)
width         | Width of a mussel (mm)
height        | Height of a mussel (mm)
size          | Mass of a mussel (g)
weight        | Weight of eatable part of a mussel (g)

We want to study how the eatable part of a mussel varies as a function of the other four variables using a multiple linear regression.

```{r mussels}

```
